---
title: "Automated Machine Learning with H2O (II)"
author: "Arash Amiririgi"
---

**Challenge:**
The goal here is to predict whether or not a product will be put on backorder status, given a number of product metrics such as current inventory, transit time, demand forecasts and prior sales. It’s a classic Binary Classification problem.

*Code:*

### Loading libraries
```{r message=FALSE, warning=FALSE}
library(tidymodels)
library(magrittr)
library(dplyr)
library(sjmisc)
library(magrittr)
library(haven)
library(sjlabelled)
library(rsample)
library(recipes)
library(rstanarm)
library(broom.mixed)
library(h2o)
```

### Loading the training & test dataset
```{r message=FALSE, warning=FALSE}
product_backorders_tbl <- read.csv("product_backorders.csv")
split_obj <- initial_split(product_backorders_tbl, prop = 0.8)
train_readable_tbl <- training(split_obj)
test_readable_tbl  <- testing(split_obj)
```

### Specifying the response and predictor variables
```{r message=FALSE, warning=FALSE}
factor_name <- "went_on_backorder"
recipe_obj <- 
  recipe(went_on_backorder ~ ., data = train_readable_tbl) %>%
  step_zv(all_predictors()) %>%
  step_mutate_at(factor_name, fn = as.factor) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  prep()
train_tbl <- bake(recipe_obj, new_data = train_readable_tbl)
test_tbl  <- bake(recipe_obj, new_data = test_readable_tbl)

h2o.init()
split_h2o <- h2o.splitFrame(as.h2o(train_tbl), ratios = c(0.8), seed = 1234)
train_h2o <- split_h2o[[1]]
valid_h2o <- split_h2o[[2]]
test_h2o  <- as.h2o(test_tbl)
y <- "went_on_backorder"
x <- setdiff(names(train_h2o), y)
```

### running AutoML specifying the stopping criterion
```{r message=FALSE, warning=FALSE}
automl_models_h2o <- h2o.automl(
  x = x,
  y = y,
  training_frame    = train_h2o,
  validation_frame  = valid_h2o,
  leaderboard_frame = test_h2o,
  max_runtime_secs  = 60,
  nfolds            = 5)
```

### View the leaderboard
```{r message=FALSE, warning=FALSE}
typeof(automl_models_h2o)
slotNames(automl_models_h2o)
automl_models_h2o@leaderboard 
automl_models_h2o@leader

extract_h2o_model_name_by_position <- function(h2o_leaderboard,
                                                n= 1, verbose = T){
   model_name <- h2o_leaderboard %>%
     as_tibble() %>%
     slice(n) %>%
     pull(model_id)
   if (verbose) message(model_name)
   return(model_name)}
```

### Predicting using Leader Model
```{r message=FALSE, warning=FALSE}
# StackedEnsemble_BestOfFamily_1_AutoML_9_20230614_11147
stacked_ensemble_h2o <- automl_models_h2o@leaderboard %>% 
  extract_h2o_model_name_by_position(7) %>% 
  h2o.getModel()

predictions <- h2o.predict(stacked_ensemble_h2o, newdata = as.h2o(test_tbl))
typeof(predictions)
predictions_tbl <- predictions %>% as_tibble()
predictions_tbl
```

### Save the leader model
```{r message=FALSE, warning=FALSE}
# 
automl_models_h2o@leaderboard %>% 
  extract_h2o_model_name_by_position(20) %>% 
  h2o.getModel() %>%
  h2o.saveModel(path = "C:\\Users\\arash\\Documents\\GitHub\\ss23-bdml-ArashAmiririgi\\content\\01_journal\\04_Modeling")
```
