---
title: "Performance Measures"
author: "Arash Amiririgi"
---

**Challenge:**
The goal here is to predict whether or not a product will be put on backorder status, given a number of product metrics such as current inventory, transit time, demand forecasts and prior sales. It’s a classic Binary Classification problem.

*Code:*

### Preparing the model based previous chapter
```{r message=FALSE, warning=FALSE}
# Loading libraries
library(tidymodels)
library(magrittr)
library(dplyr)
library(sjmisc)
library(magrittr)
library(haven)
library(sjlabelled)
library(rsample)
library(recipes)
library(rstanarm)
library(stringr)
library(broom.mixed)
library(h2o)

# Loading the training & test dataset
product_backorders_tbl <- read.csv("product_backorders.csv")
split_obj <- initial_split(product_backorders_tbl, prop = 0.8)
train_readable_tbl <- training(split_obj)
test_readable_tbl  <- testing(split_obj)

# Specifying the response and predictor variables
factor_name <- "went_on_backorder"
recipe_obj <- 
  recipe(went_on_backorder ~ ., data = train_readable_tbl) %>%
  step_zv(all_predictors()) %>%
  step_mutate_at(factor_name, fn = as.factor) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  prep()
train_tbl <- bake(recipe_obj, new_data = train_readable_tbl)
test_tbl  <- bake(recipe_obj, new_data = test_readable_tbl)

h2o.init()
split_h2o <- h2o.splitFrame(as.h2o(train_tbl), ratios = c(0.8), seed = 1234)
train_h2o <- split_h2o[[1]]
valid_h2o <- split_h2o[[2]]
test_h2o  <- as.h2o(test_tbl)
y <- "went_on_backorder"
x <- setdiff(names(train_h2o), y)

# running AutoML specifying the stopping criterion
automl_models_h2o <- h2o.automl(
  x = x,
  y = y,
  training_frame    = train_h2o,
  validation_frame  = valid_h2o,
  leaderboard_frame = test_h2o,
  max_runtime_secs  = 60,
  nfolds            = 5)

# View the leaderboard
typeof(automl_models_h2o)
slotNames(automl_models_h2o)
automl_models_h2o@leaderboard 
automl_models_h2o@leader
```

### Leaderboard vizualization

```{r}
automl_models_h2o@leaderboard %>% 
              as_tibble() %>% 
              select(-c(mean_per_class_error, rmse, mse))
```


```{r message=FALSE, warning=FALSE}
plot_h2o_leaderboard <- function(h2o_leaderboard, order_by = c("auc", "logloss"), 
                                 n_max = 20, size = 4, include_lbl = TRUE) {
    order_by <- tolower(order_by[[1]])
    leaderboard_tbl <- h2o_leaderboard %>%
        as.tibble() %>%
        select(-c(aucpr, mean_per_class_error, rmse, mse)) %>% 
        mutate(model_type = str_extract(model_id, "[^_]+")) %>%
        rownames_to_column(var = "rowname") %>%
        mutate(model_id = paste0(rowname, ". ", model_id) %>% as.factor())
    if (order_by == "auc") {
        data_transformed_tbl <- leaderboard_tbl %>%
            slice(1:n_max) %>%
            mutate(
                model_id   = as_factor(model_id) %>% reorder(auc),
                model_type = as.factor(model_type)) %>%
                pivot_longer(cols = -c(model_id, model_type, rowname), 
                       names_to = "key", 
                       values_to = "value", 
                       names_transform = list(key = forcats::fct_inorder))
    } else if (order_by == "logloss") {
        data_transformed_tbl <- leaderboard_tbl %>%
            slice(1:n_max) %>%
            mutate(
                model_id   = as_factor(model_id) %>% reorder(logloss) %>% fct_rev(),
                model_type = as.factor(model_type)
            ) %>%
            pivot_longer(cols = -c(model_id, model_type, rowname), 
                       names_to = "key", 
                       values_to = "value", 
                       names_transform = list(key = forcats::fct_inorder))
    } else {
        stop(paste0("order_by = '", order_by, "' is not a permitted option."))
    }
    g <- data_transformed_tbl %>%
        ggplot(aes(value, model_id, color = model_type)) +
        geom_point(size = size) +
        facet_wrap(~ key, scales = "free_x") +
        labs(title = "Leaderboard Metrics",
             subtitle = paste0("Ordered by: ", toupper(order_by)),
             y = "Model Postion, Model ID", x = "")

    if (include_lbl) g <- g + geom_label(aes(label = round(value, 2), 
                                             hjust = "inward"))

    return(g)
}

plot_h2o_leaderboard(automl_models_h2o@leaderboard)
```

### Tune a model with grid search
```{r message=FALSE, warning=FALSE}
deeplearning_h2o <- h2o.loadModel("04_Modeling/DeepLearning_grid_1_AutoML_9_20230614_11147_model_1")
h2o.performance(deeplearning_h2o, newdata = as.h2o(test_tbl))

deeplearning_grid_01 <- h2o.grid(
    algorithm = "deeplearning",
    grid_id = "deeplearning_grid_01",
    x = x, y = y,
    training_frame   = train_h2o,
    validation_frame = valid_h2o,
    nfolds = 5,
    hyper_params = list(
        hidden = list(c(10, 10, 10), c(50, 20, 10), c(20, 20, 20)),
        epochs = c(10, 50, 100)))

deeplearning_grid_01
h2o.getGrid(grid_id = "deeplearning_grid_01", sort_by = "auc", decreasing = TRUE)
deeplearning_grid_01_model_1 <- h2o.getModel("deeplearning_grid_01_model_1")

deeplearning_grid_01_model_1 %>% h2o.auc(train = T, valid = T, xval = T)

deeplearning_grid_01_model_1 %>%
    h2o.performance(newdata = as.h2o(test_tbl))
```

### Visualizing the trade of between the precision and the recall and the optimal threshold
```{r message=FALSE, warning=FALSE}
stacked_ensemble_h2o <- h2o.loadModel("04_Modeling/StackedEnsemble_AllModels_1_AutoML_9_20230614_11147")
deeplearning_h2o     <- h2o.loadModel("04_Modeling/DeepLearning_grid_1_AutoML_9_20230614_11147_model_1")
glm_h2o              <- h2o.loadModel("04_Modeling/GLM_1_AutoML_9_20230614_11147")
```

### ROC Plot
```{r message=FALSE, warning=FALSE}

```

### Precision vs Recall Plot
```{r message=FALSE, warning=FALSE}

```

### Gain Plot
```{r message=FALSE, warning=FALSE}

```

### Lift Plot
```{r message=FALSE, warning=FALSE}

```

### Dashboard with cowplot
```{r message=FALSE, warning=FALSE}

```
